\documentclass[11pt,a4paper,twoside]{article}
\usepackage{mha}
\begin{document}
\MHAtitle{The auditory profile library 'libaudprof'}
\newpage
\MHAcopyright{}
\newpage
\tableofcontents
\newpage
\renewcommand{\leftmark}{\rightmark}
\pagenumbering{arabic}

\section{Introduction}

The MHA uses the concept of auditory profile based fitting. An
auditory profile is a collection of clinical data related to hearing
disabilities. Components which are used at time of writing are air and
bone conduction hearing thresholds, uncomfortable threshold and
categorical loudness scaling data. Although most fitting rules utilize
only the hearing threshold, and only few use the loudness scaling
data, the underlying library functions offer ways for extension to
more parameters.

Typically, the fitting is controlled via the graphical interface
\verb!mhacontrol!. This document decribes solutions for selecting
auditory profiles from the data base manually and to upload fittings
to an MHA instance.

\section{Accessing the auditory profile database}

Auditory profiles in the MHA are stored in MATLAB format in a file
called \verb!audprofdb.mat!. A graphical interface to this data base
is the MATLAB program \verb!clientdb!.

To access the auditory profile data base directly, the MATLAB library
\verb!libaudprof! can be used, to return a structure with function
handles:
\begin{verbatim}
h = libaudprof();
\end{verbatim}
To get help to any sub function, e.g., \verb!db_load!, the help
function can be called with:
\begin{verbatim}
h.help.db_load()
\end{verbatim}
The function to load the auditory profile 'slope' from the client with
ID 'TT123456' is
\begin{verbatim}
cDB = h.db_load();
sClientID = 'TT123456';
sAudID = 'slope';
sAud = h.audprof_get( sClientID, sAudID, cDB );
\end{verbatim}
A shortcut for loading all auditory profiles, and a solution to load
only the last one in list, is
\begin{verbatim}
csAud = h.audprof_getall( sClientID );
sAud = csAud{end};
\end{verbatim}
Here, the data base is automatically loaded.

To import air conduction thresholds from an XLS sheet, one can use
\begin{verbatim}
xlsname = 'myaud.xls';
h.db_xlsimport( xlsname );
% verify the imported data:
clientdb();
\end{verbatim}
The xls file needs to follow a special formatting: The first row is
interpreted as a header row. The field A1 may contain any text, the
fields B1, C1, D1, etc. need to specify the side and frequency in a
string, e.g., ``l125''. The fields are not case sensitiv and may
contain spaces. The frequencies must be given in Hz. The side
identifier can be 'l' or 'r'. The first column must contain the client
identifier, which are converted into upper case during the import
process. The auditory profile identifiers are set to the xls file
name. Auditory profiles of the same identifier and client ID, which
existed before in the database, are overwritten.

To fit a running MHA (either in 'running' or in 'prepared' state, see
user manual for details), one function from the \verb!libmultifit!
library can be used:
\begin{verbatim}
hMF = libmultifit();
sGainrule = 'NALRP';
sSide = 'lr';
csPlugs = hMF.uploadallfirstfit( mha, sGainrule, sAud, sSide );
\end{verbatim}
For bilateral fittings, the side string must be 'lr', otherwise the
side to be fitted should be selected ('l' or 'r').

If an auditory profile contains any Inf or NaN values, they should be
removed before creating hearing aid fittings:
\begin{verbatim}
sAud = h.audprof_cleanup( sAud );
\end{verbatim}

\bibliography{MHA}

\printindex

\end{document}

% LocalWords:  audiogram MHA LTASS
