\section{The MHA Kernel}
\label{sec:kernel}

The MHA kernel is responsible for the signal flow management in the MHA and
for the communication between MHA plugins (processing blocks). The kernel is
loaded and controlled by the framework application. Processing plugins can be
inserted into processing chains. Multiple chains can be configured if
switching between complex configurations is required. The inactive processing
chains can optionally be calculated to reduce switching noise or for pure
signal analysis. Time domain (waveform) and frequency domain (STFT) processing
is supported. The MHA plugins can process in a single domain or as a domain
transformation (e.g. FFT).

The MHA kernel provides a multichannel environment. Any number of input and
output channels can be configured (although some standard hearing aid
algorithms make sense only with two channels).
%
Down-mixing and up-mixing in the plugins is possible.
%
This is highly relevant for processing of microphone arrays.
%
At the moment, multichannel processing is available in the \mhad{}
only; the \Windows{} version is restricted to two channels.
%
The \mhad{} optionally provides monitoring of all signal processing
chain outputs.

\subsection{Configuration variables of the MHA Kernel}

\begin{description}
  \mhavardesc{disable\_inactive}{Disable inactive chains. If set to {\tt yes}
    then only the selected chain is calculated.}
  
  \mhavardesc{chains}{The list of signal processing chains which will be
    configured. Each entry will appear as a sub-parser which holds the chain
    specific configuration including the configuration of the algorithms.}
  
  \mhavardesc{chain}{Name of the selected chain. Setting this variable will
    select one of the previously configured chains.}
    
  \mhavardesc{chainname.inputdomain}{The domain which will be passed to
    the first plugin in a chain ({\tt waveform} or {\tt spectrum}).}
  
  \mhavardesc{chainname.algos}{Assuming that {\tt chainname} is an entry of
    the variable {\tt chains}, then this is a list of algorithms to run in
    this chain. Each entry of this variable is the basename of a MHA Plugin.
    The plugin configuration will appear as a sub-parser of {\tt chainname}.
    The plugin basename is used as a sub-parser name except when explicitly
    specified by typing {\tt basename:parsername}. This is especially
    relevant when using multiple instances of the same algorithm in one chain.
    The order of algorithms in this list determines the processing order (from
    left to right).}

\mhavardesc{fftname}{name of the FFT library without extension (e.g.\
  {\tt wave2spec}).}

\mhavardesc{ifftname}{name of the inverse FFT library without
  extension (e.g.\ {\tt spec2wave}).}
  
\mhavardesc{fft}{subparser containing the configuration of the FFT
  plugin, see section \ref{sec:wave2spec} for details.}

\mhavardesc{ifft}{subparser containing the configuration of the
  inverse FFT plugin, see section \ref{sec:spec2wave} for details.}
\end{description}

%% \subsection{Monitor variables of the MHA Kernel (profiling)}
%% 
%% The MHA Kernel provides a system to watch the time spend in the
%% configured chains and in each algorithm/plugin processing
%% callback. This profiling method can be used to roughly estimate the
%% relative processing power of an algorithm if the system is not loaded
%% by other processes. These monitor variables are available:
%% 
%% \begin{description}
%% \mhavardesc{mhatime}{This variable holds the run time in seconds since
%%   the last call of the prepare function. If the framework is prepared
%%   just before entering the running state this variable is comparable
%%   to the real running time (real time).}
%% 
%% \mhavardesc{mhaprof}{The cumulative time used by the MHA Kernel
%%   processing callback in seconds. This includes the time spent in the
%%   MHA Plugins.}
%% 
%% \mhavardesc{chainprof}{The cumulative processing time used by the
%%   chains in seconds. Each entry is the time used by one chain. See the
%%   {\tt chains} variable for the corresponding chain names.}
%% 
%% \mhavardesc{chainname.algoprof}{The cumulative processing time used by
%%   the algorithms of one chain of name {\tt chainname}. Each entry is
%%   the time used by one algorithm (MHA Plugin). See the {\tt algos}
%%   variable for the corresponding plugin names.}
%% 
%% \mhavardesc{prof.mha}{Cumulative processing time in seconds of MHA
%%   Kernel.}
%% 
%% \mhavardesc{prof.fft}{Cumulative processing time in seconds of FFT
%%   library.}
%% 
%% \mhavardesc{prof.ifft}{Cumulative processing time in seconds of
%%   inverse FFT library.}
%% \end{description}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "MHA_manual"
%%% End: 
