\section{Hints and links for tuning the realtime environment}%
\label{sec:system}%

Low delay real-time signal processing is a task which depends highly on
the operation system performance. For low latency audio processing
with a total delay of 4-6 ms, the maximal system latency needs to be
as low as 1 or 2 ms.  On a single processor high level operating
system (e.g. Linux, \Windows{}), multitasking is usually reached by
sequentially processing each task only for a limited period of time
and than switching to the next task. This method is obviously not
suitable for real-time signal processing since the execution of code
can be delayed by an unpredictable amount of time. Low latency
real-time processing tasks therefore have to be started in a special
mode which grants the execution of its code. Furthermore, the system
has to be manipulated in way which reduces the maximal interruption
time by low level system tasks (e.g. accesses to hard disks or graphic
cards) or reduces their priority below the priority of the real-time
process. For Linux operating systems, a modified kernel is available
which provides these features.

\subsection{Linux audio distributions}

To manually patch a Linux kernel and configure the operating system
for optimal audio processing is a long and difficult task. We rather
recommend to use a Linux distribution which is prepared and optimised
for audio processing. At least two audio distributions are freely
available, one of those is used by H\"orTech for low delay audio
processing.

A widely used audio distribution is 'Planet CCRMA'\index{audio distribution}
which is built on top of a RedHat or Fedora Linux distribution. It is easy to
maintain and includes all software packages required for low latency\index{low
  latency} signal processing with the MHA (mainly the JACK\index{JACK} sound
server, ALSA\index{ALSA} sound card drivers and a low latency kernel). System
updates including security fixes are available. This distribution is actively
maintained. This distribution is used by H\"orTech.  More information and
download sites can be found here:

\url{http://ccrma.stanford.edu/planetccrma/software/}

Another audio distribution developed and compiled as an IST project
is 'AGNULA'. Two versions of this distribution exists, one built on
top of a Debian distribution, the other on top of RedHat Linux. This
audio distribution contains all packages required to run the low
latency realtime MHA (JACK sound server, ALSA sound card drivers and a
low latency kernel). Watching the project in the last months it seems
that the maintenance is only performed at a low rate. H\"orTech has no
direct experience with this distribution. More information and
download sites can be found following this link:

\url{http://www.agnula.org/}

Both distributions provide a detailed discussion on low latency signal
processing in general and on system tuning. We strongly recommend to
have a look at these sites before starting to work with low latency
processing. Another source of information about low latency audio
processing can be found here:

\url{http://www.Linuxdj.com/audio/quality/}

Information on the ALSA (Advanced Linux Sound Architecture) sound
drivers and supported audio devices can be found in the web:

\url{http://www.alsa-project.org/}


\subsection{The JACK low latency sound server}
\label{sec:jack}

`JACK is a low-latency audio server, written for POSIX-conforming
operating systems such as GNU/Linux and Apple's OS X. It can connect a
number of different applications to an audio device, as well as
allowing them to share audio between themselves. Its clients can run
in their own processes (ie. as normal applications), or they can run
within the JACK server (ie. as a plugin).'

`JACK was designed from the ground up for professional audio work, and
its design focuses on two key areas: synchronous execution of all
clients, and low latency operation.' (citation from the JACK web site).

The \mhad{} can use the JACK\index{JACK} low latency sound
server for audio input and output.
%
The advantage of using JACK in opposite to directly using the sound
driver layer IO is the possibility to connect to many (almost any)
audio clients.
%
At the same time it passes low latency features of the driver layer
(namely ALSA, but other drivers are supported as well) to the client.
%
JACK is available for Linux and Mac OS X. Documentation and download
sites can be found at this address:

\url{http://www.jackaudio.org/}

A JACK client can be added to a running sound server. While the client
is active, it can be connected to other clients or hardware ports
through API functions, command line tools or graphical user
interfaces. Multiple connections to or from a client port are
possible. Links to many useful tools, e.g. mixing
tools, graphical control interfaces, signal analysis tools, can be
found on the JACK website.

\subsubsection{Invocation of JACK}

The JACK\index{JACK} sound server has to be started before the MHA. Once
started, the configuration (fragment size, sampling rate) of JACK is fixed. To
change these parameters, please close all JACK clients and restart the server.
Details on the invocation of JACK are given in the \verb!jackd!  manual page
and in the package documentation. Here only MHA specific items will be
discussed.

Best performance will be reached if JACK uses direct hardware access
with its native parameters. Usually this is provided by the ALSA 'hw'
device. Sometimes it is necessary to add device and subdevice number
to the device name, e.g. in case of an RME Digi 96 configured as the
second sound card use \verb!hw:1,1! to address its eight channel ADAT
mode. When using the hardware device \verb!hw!, only native parameters
are supported. This means that only a restricted set of fragment sizes
(JACK: \verb!--period!, \verb!-p!), number of hardware buffers (JACK:
\verb!--nperiods!, \verb!-n!) and sampling rates (JACK: \verb!--rate!,
\verb!-r!)  can be configured.

If it is required to use non-native sampling rate some problems may
occur. Due to buffer size restrictions (JACK allows only powers of
two, ALSA requires the ratio between native and user sampling rate to be
the same as the ratio between hardware period size and user period
size) only down-sampling by a power of two is supported with the ALSA
plugin driver. Therefore a sampling rate of 16 kHz can only be reached
when using sound cards which support 32 kHz or 64 kHz sampling rate
(e.g. not supported by the ALSA driver for RME Digi 96). The following
entry might be needed in your \verb!~/.asoundrc! file in order to
work properly with your card:
\begin{verbatim}
pcm.mhadev {
    type plug
    slave {
        pcm "hw:1,1"
        rate 32000
    }
}
\end{verbatim}
Please replace \verb!hw:1,1! by the correct device name of your sound
card. The JACK daemon now can be started using the \verb!mhadev! sound
device:
\begin{verbatim}
jackd -d alsa -d mhadev -r 16000 -p 128 -n 2
\end{verbatim}
This will use the sound card \verb!hw:1,1! with the native sampling
rate 32 kHz, a hardware buffer length of 256 samples and two hardware
buffers. Warning messages about using the ALSA software "plug" layer 
will be shown when not using the hardware device \verb!hw!.

However, if all this doesn't work it is still possible to use the OSS
driver interface of JACK for sound card access. With OSS it should be
possible to configure non-native sampling rates more easily, with the
disadvantage of possibly working with longer delays and without direct
control of the audio hardware parameters.

If a JACK plugin for ALSA is installed (e.g.\ included in the Planet
CCRMA distribution), it might be useful to define a virtual ALSA
device, which automatically connects to the MHA JACK client:
\begin{verbatim}
pcm.mha {
    type plug
    slave { 
        pcm {
            type jack
            playback_ports {
                0 MHA:in_1
                1 MHA:in_2
            }
            capture_ports {
                0 MHA:out_1
                1 MHA:out_2
            }
        }
    }
}
\end{verbatim}
When JACK and the MHA are running, a sound file can be played through
the MHA by typing:
\begin{verbatim}
aplay -D plug:mha soundfile.wav
\end{verbatim}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "openMHA_application_manual"
%%% End: 
