%%% This file is part of the Open HörTech Master Hearing Aid (openMHA)
%%% Copyright © 2018 2019 2020 HörTech gGmbH
%%%
%%% openMHA is free software: you can redistribute it and/or modify
%%% it under the terms of the GNU Affero General Public License as published by
%%% the Free Software Foundation, version 3 of the License.
%%%
%%% openMHA is distributed in the hope that it will be useful,
%%% but WITHOUT ANY WARRANTY; without even the implied warranty of
%%% MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
%%% GNU Affero General Public License, version 3 for more details.
%%%
%%% You should have received a copy of the GNU Affero General Public License, 
%%% version 3 along with openMHA.  If not, see <http://www.gnu.org/licenses/>.

% Latex header for doxygen 1.8.11
\documentclass[11pt,a4paper,twoside]{article}

% Packages required by doxygen
\usepackage{fixltx2e}
\usepackage{calc}
\usepackage{../openMHAdoxygen}
\setlength{\headheight}{13.6pt}
\usepackage[export]{adjustbox} % also loads graphicx
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{makeidx}
\usepackage{multicol}
\usepackage{multirow}
\PassOptionsToPackage{warn}{textcomp}
\usepackage{textcomp}
\usepackage[nointegrals]{wasysym}
\usepackage[table]{xcolor}

% Font selection
\usepackage[T1]{fontenc}
\usepackage{helvet}
\usepackage{courier}
\usepackage{amssymb}
\usepackage{sectsty}
\renewcommand{\familydefault}{\sfdefault}
\allsectionsfont{%
  \fontseries{bc}\selectfont%
  \color{darkgray}%
}
\renewcommand{\DoxyLabelFont}{%
  \fontseries{bc}\selectfont%
  \color{darkgray}%
}
\newcommand{\+}{\discretionary{\mbox{\scriptsize$\hookleftarrow$}}{}{}}

% Page & text layout
\usepackage{geometry}
\geometry{%
  a4paper,%
  top=2.5cm,%
  bottom=2.5cm,%
  left=2.5cm,%
  right=2.5cm%
}
\tolerance=750
\hfuzz=15pt
\hbadness=750
\setlength{\emergencystretch}{15pt}
\setlength{\parindent}{0cm}
\setlength{\parskip}{3ex plus 2ex minus 2ex}
\makeatletter
\renewcommand{\paragraph}{%
  \@startsection{paragraph}{4}{0ex}{-1.0ex}{1.0ex}{%
    \normalfont\normalsize\bfseries\SS@parafont%
  }%
}
\renewcommand{\subparagraph}{%
  \@startsection{subparagraph}{5}{0ex}{-1.0ex}{1.0ex}{%
    \normalfont\normalsize\bfseries\SS@subparafont%
  }%
}
\makeatother

% Headers & footers
\usepackage{fancyhdr}
\pagestyle{fancyplain}
\renewcommand{\sectionmark}[1]{%
  \markright{\thesection\ #1}%
}
\fancyhead[LE]{\fancyplain{}{\bfseries\thepage}}
\fancyhead[CE]{\fancyplain{}{}}
\fancyhead[RE]{\fancyplain{}{\bfseries\leftmark}}
\fancyhead[LO]{\fancyplain{}{\bfseries\rightmark}}
\fancyhead[CO]{\fancyplain{}{}}
\fancyhead[RO]{\fancyplain{}{\bfseries\thepage}}
\fancyfoot[LE]{\fancyplain{}{}}
\fancyfoot[CE]{\fancyplain{}{}}
\fancyfoot[RE]{\fancyplain{}{\bfseries\scriptsize \copyright{} 2005-2020 H\"orTech gGmbH, Oldenburg }}
\fancyfoot[LO]{\fancyplain{}{\bfseries\scriptsize \copyright{} 2005-2020 H\"orTech gGmbH, Oldenburg }}
\fancyfoot[CO]{\fancyplain{}{}}
\fancyfoot[RO]{\fancyplain{}{}}

% Indices & bibliography
\usepackage{natbib}
\usepackage{tocloft}
\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{4}
\addtolength{\cftsubsecnumwidth}{5pt}
\makeindex

% Custom commands
\newcommand{\clearemptydoublepage}{%
  \newpage{\pagestyle{empty}\cleardoublepage}%
}

\usepackage{caption}
\captionsetup{labelsep=space,justification=centering,font={bf},singlelinecheck=off,skip=4pt,position=top}

%===== C O N T E N T S =====

\begin{document}

\MHAtitle{The \mha{} Calibration Manual}
\newpage
\MHAcopyright{}
\newpage
\tableofcontents
\newpage
\renewcommand{\leftmark}{\rightmark}
\pagenumbering{arabic}

\section{Introduction}

The H\"{o}rtech Open Master Hearing Aid (\mha{}) is a software platform
for algorithm development and evaluation.
%
For certain aspects of algorithm evaluation, correct calibration of the
test setup may be a requirement.
%
Since \mha{}, being a software platform, does not include hardware
(like hearing aid housings, microphones, and receivers),
users of the \mha{} are expected to supplement suitable hardware
to pick up and produce sound at the users' ears.
%
Users then face the challenge of how to correctly configure the software
to calibrate the system.

This guide is a collection of best practices that the authors and users
of \mha{} employ to ensure their setups are calibrated correctly.
%
We will extend this guide with new procedures as we add them to the
openMHA, and as we receive user contributions in the field of
calibration.
%
\mha{} users should be aware that calibration is an involved topic.
%
Users need professional, specialized acoustic measurment equipment in
order to be able to perform accurate measurements, and understand
how to use their equipment.
%
Understanding how to calibrate commercial hearing aids is a requirement
for using this guide.
%
The book "Hearing Aids" by Harvey Dillon contains a suitable treatment
of the topic in chapter 4 of its second edition (2012).
%
We cannot guarantee that the procedures presented here are free of
errors.
%
If you discover something that is wrong, please get in contact with
the \mha{} project, e.g. by filing an issue with our github project.
%
We will do our best to improve the documentation and the tools that we
provide as we go along.

\section{Calibrating hearing aid outputs}

This section describes how to measure the frequency response of a
custom hearing aid output hardware, and how to use this frequency
responses to calibrate the output of the device.

%

\subsection{Considerations regarding calibration of custom hardware used as hearing aid outputs}

Hearing aid receivers typically have a low impedance and a non-flat frequency response.
%
When attaching hearing aid receivers to the output of a computer's
sound card, care has to be taken to ensure that the sound card can
actually drive the receiver, i.e.\ provide the necessary current
without voltage drops.
%
Otherwise serious non-linearities will result.
%
Consider the application of a suitable amplifier to drive your output hardware.

A suitable acoustic coupler has to be used to connect the measurement
microphone to the hearing aid receiver.
%
The coupler will simulate the mechanical properties similar to
a real ear canal.
%
The coupler will also influence how the measurment microphone of your
measurement equipment has to be calibrated.
%
Please refer to Dillon and to the manuals of your measurement
equipment for details.

When you calibrate custom hearing aid hardware, you always calibrate
the complete output chain in combination, consisting of
%
the individual sound card, any amplifier that you employ, and the
hearing aid receiver.

Hearing aids need to produce ear canal sound levels, not free field
sound levels.
%
For this reason, we suggest to also apply the Real-Ear-Unaided-Gain
(REUG) in the \mha{} output calibration.
%
Please refer to Dillon for an explanation of REUG.

When calibrating hearing aid devices, you need to make a choice
whether you are ok with calibrating the device for an average ear, or
for the individual ear of a specific subject.
%
In the following text, we will describe only the average calibration
method. This may change in future.

Calibration requires reliable measurement of acoustic levels.
%
This can only be achieved in a quiet environment.
%
Please perform all acoustic measurements in a suitable, sound-proof
booth or a similar environment.
%
If your calibration equipment supports it, consider to filter out very
low frequencies (e.g. < 20 Hz).
%
If you notice fluctuations in the measured level, e.g. resulting from
doors on other levels in your building, or from traffic outside,
consider moving to a better isolated location, or to off-peak hours.
%
To ensure that the influence of unavoidable noise is minimal, acoustic
calibration is performed at relatively high sound levels around 80 dB.
%
If you cannot achieve these levels with your custom output hardware,
you should consider modifying the output hardware rather than
measuring a lower sound level.

We will be using the MHA to produce test tones of which we then
measure the acoustic levels.
%
When producing test tones with high frequencies, take into account
that the sound card will apply a low-pass filter for anti-aliasing
that affects already frequencies below the Nyquist frequency.
%
This can be an issue if you want to calibrate your system for
relatively low acoustic sampling rates: If e.g. you use 16 kHz
sampling rate in your sound card, measuring the sound level at 8 kHz
should be avoided, and you should not reuse the measurement taken at 6
kHz for higher sound card sampling rates.

The calibration process will probably run into practical limits, and
will require compromises.
%
E.g., hearing aid receivers typically produce significantly softer
sound levels for the same electric voltages at low frequencies.
%
Exact calibration requires filters that invert the frequency response
of the hearing aid output hardware, which would mean in this case,
amplify with a high gain at low frequencies.
%
However, a high enough gain at low frequencies will cause the output
signal amplitude to exceed the dynamic range of the sound card.
%
Since it is not possible to produce signal with a higher amplitude
than the maximum amplitude of the sound card, the signal would need to
be limited after the calibration filter is applied. This limiting will
cause audible artifacts.
%
To avoid these situations, it is usually a better choice to avoid too
high amplification at the boundaries of the transmitted spectrum and
to accept that the output levels are softer than intended by the
hearing aid signal processing algorithms in these frequency areas.

\subsection{Measuring the frequency response for a custom hearing aid output hardware with openMHA with an IEC 711 coupler}

We describe here how to use a coupler according to IEC 60 711 to
measure the frequency response of custom hearing aid output hardware.
%
Compared to the 2cc coupler, the IEC 711 coupler has the advantage
that for an average ear, the Real-Ear-to-Coupler-Difference (RECD) is
small enough that it does not need to be taken into account when doing
measurements to calibrate for the average ear.

Please assemble and calibrate your measurement equipment to perform
measurements with the IEC 711 coupler.
%
Please refer to the manual of the equipment manufacturer for
instructions.
%
Note that some calibrator devices require that you take correction
factors into account when calibrating your equipment for the IEC 711.

% We will talk of classes of custom hearing aid outputs in the next iteration.
%
% When we talk of a class of custom hearing aid output software, we
% refer to an ensemble of output devices that use the same type of
% receiver, the same type of amplifier, and the same type of soundcard.
% %
% Individual receivers, amplifiers, and soundcards will differ from each
% other, even if they are of the same type.
% %
% You will have to decide for yourself what difference you are going to
% accept and still consider the devices to be in the same class.

When your equipment is calibrated, then you can measure the frequency
response of the output hardware with the help of the openMHA:

\begin{itemize}
\item
  Install \mha{} on the computer where the calibration is performed (see README.md).
\item
  We will use \Octave{}/\Matlab{} tools to control the measurement. To
  prepare,
  \begin{itemize}
  \item Start \Octave{}/\Matlab{} from the shell where you have set
    the \mha{} environment variables.
  \item Change to the directory mha/tools/mfiles.
  \item Execute \texttt{javaaddpath mhactl\_java.jar} inside
    \Octave{}/\Matlab{}.
  \end{itemize}
\item
  Start the jack sound server with the same sampling rate that you
  will later use with the \mha{}.
\item
  Connect the custom hearing aid output hardware to the desired
  sound card output channel and to the measurement equipment.
\item
  Use the \Octave{}/\Matlab{} function \texttt{measure\_fresponse}
  provided with \mha{} to measure the frequency response:
  \begin{itemize}
  \item Call this function with the following parameters:
    \begin{enumerate}
    \item the name of the sound card channel in jack,
      e.g. \texttt{'system:playback\_1'},
    \item the sampling rate of the jack server in Hz.
    \end{enumerate}
  \item The function returns a struct containing the fields:
    \begin{description}
    \item [Frequencies] A vector of frequencies in Hz of the test tones,
    \item [dBFSfor80dB] A vector of corresponding levels in dB re full scale (dB FS) of the sound card where the acoustic level was 80 dB.
    \item [REUGdB] A vector of corresponding suggested correction gains in dB to compensate for REUG.
    \item [correctionsdB] A vector of corresponding correction gains, initially all 0.
    \end{description}
  \item
    example: \texttt{response = measure\_fresponse('system:playback\_1',48000);}
  \item
    The function will start an MHA and begin to output test tones.
  \item
    Your measurement equipment which is connected to the output will
    measure the acoustic level produced by each test tone.
  \item
    Your goal is to alter the output levels so that acoustic levels of
    80 dB are produced by the test tones:
  \item
    If the acoustic level is different from 80 dB, then adjust the level
    by entering the required difference to reach 80 dB.
  \item
    Repeat the corrections until your setup produces 80 dB acoustic
    output level.
  \item
    When you have reached 80 dB, enter a correction value of 0 to move
    on to the next test tone.
  \end{itemize}
\end{itemize}
  
You have now the frequency response of your setup in a struct in
\Octave{}/\Matlab{}.
%
You can use all the tools available in \Octave{}/\Matlab{} to inspect
and alter the measured frequency response in this struct before using
it to calibrate the MHA.
%
We provide the \Octave{}/\Matlab{} function \texttt{plot\_fresponse}
that you can use to plot the frequency response.
%
The plot will display the frequency response of the required filter to
calibrate the MHA for your output hardware.
%
This filter inverts the frequency response of your output hardware, and
adds the gains to apply the REUG.

You now have to decide if the overall dynamic exhibited by this filter
is small enough that you want to apply the calibration filter as it
is, fully compensating the averaged frequency response, of if you want
to introduce correction gains to avoid too high gains at high or low
frequencies, by changing the correctionsdB vector in the frequency
response struct in \Octave{}/\Matlab{}.

\subsection{Using the measured frequency response to calibrate a hearing aid output device}

The \mha{} plugin \texttt{transducers} is used to calibrate inputs and outputs.
%
\texttt{transducers} can be configured with FIR filter parameters and a peak level
to compensate for the frequency responses of hearing aid microphones and receivers.
%
To make use of this capability, choose a \mha{} configuration that already includes
the \texttt{transducers} plugin, or create a new configuration that contains
\texttt{transducers}.

When you are content with the calibration filter's frequency response,
you can convert the frequency response struct to a set of filter
coefficients and peak level that can be used in \mha{} configuration.
%
The \Octave{}/\Matlab{} function \texttt{fresponse\_to\_cfg} can be
used to convert the frequency response struct to a text string that
can be used in the \mha{} configuration.

The function \texttt{fresponse\_to\_cfg} prints out values for the
\texttt{peaklevel} and for the \texttt{fir} configuration variables of
the \mha{} plugin \texttt{transducers}.

You can take these values and insert them into your configuration file,
into the \texttt{transducers} configuration.
%
Remove the elipses (\texttt{...}) printed, and remove superfluous
semicolons.
%
These are printed by the \texttt{fresponse\_to\_cfg} for cases where
you have multiple output channels. You have to assign the peaklevels
and FIR filter coefficients to the correct output channels that
correspond to the measured hardware.

\subsection{Test your calibration}

It is crucial to test your calibration. Please configure a simple \mha{} with the
calibration values that you have measured, insert the sine plugin into this \mha{},
and play back sinusoids with different frequencies and different levels.
%
A suitable configuration file is present in the \texttt{examples} directory of \mha{}.
%
Please use the configuration file in the \texttt{test-calibration} subdirectory
and alter it to contain your specific calibration values.
%
The configuration can then be started with
\texttt{mha ?read:test-calibration.cfg cmd=start}.
%
You can either modify the sinusoid frequencies in this configuration
file and start \mha{} repeatedly with different values of
\texttt{mha.sine.f}, or use the network or \Octave{}/\Matlab{} based
tools to modify this value in the running \mha{}.

When playing back sinusoids at the measurement frequencies, check the output levels with your
measurement equipment. It is now expected that the output levels for 80 dB sinusoids is measured as 80 dB + REUG(f) + correction(f).

Differences of up to a few dB are expected. You should check that over
the total spectrum, the differences are equally distributed to higher
and lower output levels than expected.

\bibliography{MHA}

\printindex

\end{document}

% LocalWords:  audiogram \mha{} LTASS

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "openMHA_application_manual"
%%% indent-tabs-mode: nil
%%% coding: utf-8-unix
%%% End:
